#Stroke Prediction - Team 30 Group Project - MGT 6203 Edx Summer 2022
#Stroke Prediction Data
#Data can be downloaded from https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset

```{r}
sh <- suppressPackageStartupMessages
sh(library("dplyr"))
sh(library("mice"))
sh(library("tidyr"))
sh(library("ggplot2"))
sh(library("Hmisc"))
sh(library("outliers"))
sh(library("naniar"))

```

#*******************************************************************************************
#Section 1 : Data Exploration and Cleaning
#*******************************************************************************************

#1.1.Loading the data and overall review of the variables 

```{r}
df <- read.table("healthcare-dataset-stroke-data.csv", sep = ",", header = T)
df <- df[,-1]
#df$bmi <- as.numeric(as.character(df$bmi))
Hmisc::describe(df)
#table(duplicated(df))
```

#The first column of the data, "id", had been eliminated for the analysis as it doesn't bear any meaningful relation to "stroke" or other data variables. The data doesn't have any duplicated rows, which is good.

#Independent Variables:

#"gender" - character with no missing values
#"age" - numeric with no missing values
#"hypertension" - numeric with no missing values
#"heart_disease" - numeric with no missing values
#"ever_married" - character with no missing values
#"work_type" - character with no missing values
#"Residence_type" - character with no missing values
#"avg_glucose_level" - numeric with no missing values
#"bmi" - character with 201 NA
#"smoking_status" - character with no missing values

#Dependent Variable:

#"stroke" - numeric with no missing values

#1.2.Converting the bmi data type from character to numeric. Since the number of NAs in bmi is 201 which is 3.9% < 5% of the whole data set, it was decided to use mean to fill in the NAs.

```{r message=FALSE, warning=FALSE, include=FALSE}
df$bmi <- as.numeric(as.character(df$bmi))
m_mice <- mice(df, m = 5, method = "mean")
df <- complete(m_mice)
```

#1.3.For better data presentation and analysis purposes, the "heart_disease" and "hypertension" were change to factors of "Yes" or "No" instead of "1" and "0". This may be changed depending on the type of data analysis later in this code. 

```{r}
df$heart_disease <- factor(df$heart_disease, levels = c(0,1), labels =c('No','Yes'))
df$hypertension <- factor(df$hypertension,levels = c(0,1), labels = c('No','Yes'))
```

#1.4. The "gender" variable contains 2994 "Female", 2115 "Male", and 1 "Other" categories. Since the "other" cateogry is only 1 in this data set, it is best to remove it ro avoid misanalysis. 

```{r}
df[df$gender=='Other', ] <- NA
df <- na.omit(df)
```

#1.5. Acknowledging the imbalances in the data set for each variable. 

```{r}
#hist(df$avg_glucose_level,xlab = 'Glucose level')
#hist(df$age,xlab = 'Age')
#hist(df$bmi,xlab = 'BMI')
#table(df$stroke)
```


#Independent Variables:

#"gender" - Imbalanced with 59% "Female" and 41% "Male".
#"age" - There is enough data with an acceptable balance before 40 and after 40
#"hypertension" - Approximately 90% "No" and 10% "Yes" - Significantly imbalanced 
#"heart_disease" - 95% "No" and 5% "Yes" - Significantly imbalanced 
#"ever_married" - Imbalanced with approximately 65% "Yes" and 35% "No".

#"work_type" - Five various categories 
#Value           children      Govt_job   Never_worked       Private   Self-employed
#Proportion         14%           13%         <1%              57%           16%

#"Residence_type" - 50% balanced between "rural" and "urban"
#"avg_glucose_level" - Major heteroscedasticity can be observed from the histogram
#"bmi" - Minor heteroscedasticity can be observed from the histogram 

#"smoking_status" - character with no missing values
#Value      formerly smoked    never smoked          smokes         Unknown
#Proportion       17%              38%                15%             30%

#Dependent Variable:

#"stroke" - 4.3% indicating stroke and 95.7% with no stroke occurance 

##1.6. Looking for outliers. The "bmi" found to have some outliers. The four outliers above 70 were removed from the data set. 

```{r}
#gt <- grubbs.test(df[,9], type = 11, opposite = FALSE, two.sided = FALSE)
#cat(gt$alternative)
boxplot(df$bmi)
df <- df %>% slice_min(bmi, n = nrow(df)-4)
```

##1.7. Imputing the the data that are not recorded for smoking. Approximately 30% of the "smoking_status" variable is recorded as "unknown" which is a significant portion. In order to reduce this value, the "smoking_status" for the ages equal or below 7 were converted to "never smoked". This reduced the 30% "uknown"smoking_status to 22%. No further assumption is made at this point, and dealing with smoking is left for each method. 

```{r}
df[df$age <= 7, "smoking_status"] <- "never smoked"
```

##1.8. The data has been randomly divided to 60% for training, 20% for validation and 20% for testing. 

#**IMPORTNAT NOTES TO REMEMBER FOR EACH METHOD"**

#The "stroke" == 1 is comprises only 5% of the data and cost analysis is required for each model.
#The "smoking_status" has 22% "uknown" catergory data points that may or may not be used in each model
#There is an imbalance between the Female and Male population. 
#There is heteroscedasticity for "bmi" and "avg_glucose_level"
#DON'T FORGET SCALLING, AS AND WHEN REQUIRED. 

#**                                            **

```{r}
set.seed(123)

N <- sort(sample(nrow(df),round(nrow(df)*0.6)))
df_train <- df [N,]
df_valtest <- df [-N,]
M <- sort(sample(nrow(df_valtest),round(nrow(df_valtest)*0.5)))
df_val <- df_valtest [M,]
df_test <- df_valtest [-M,]


```

#*******************************************************************************************
#Section 2 : Data Classification and Prediction
#*******************************************************************************************

#ATTENSION TO GROUP: Don't forget to use "reg_" or "svp-" or "knn_" or "tree_" before everything you define depending in which model you use. I even suggest you to creat df_train, df_test, and df_val with these prefixes for your model before working with them. This will prevent from duplicates and conflicts of any sort. Thanks! 